On Determinism of Game Engines used for Simulation-based Autonomous Vehicle Verification

### 2020-11-13 paper/cav_determinisim_main.tex
Modified methodology diagram
Reorganised some bits in the intro

\subsection{Definition: Determinism}
% What is a deterministic system
Schumann describes determinism as the property of causality given a temporal development of events such that any state is completely determined by prior states~\cite{Schumann2010}. However, in the context of simulation this should be expanded to include not just prior states but also the history of actions taken by all actors. Therefore, a deterministic simulation will always produce the same result given the same history of prior states and actions.
% 
% Herein the term \textit{simulation variance} is used to refer to how simulation traces and outputs vary, when the same stimulus is provided repeatedly to the simulation environment under identical initial conditions. Specifically, we focus on the variance of actor paths within a given scenario.

A simulation can be thought of as the generation or production of experimental data. 
%
In the case of a driving simulator, kinematics will describe future states of actors given the current conditions and actions taken, thereby generating new data. 
%
If a simulation is deterministic then there will be no variation in the generated output data, i.e. all future states are perfectly reproducible from prior states and actions. 
%
However, if a simulation is non-deterministic then then there will be a variation in the generated output data.

@@

\subsection{Definitions: Precision, Variance \& Tolerance}

@@

If a simulator is deterministic then it will produce results with absolute precision or zero variance and hence will be within acceptable tolerance. If the simulator is non-deterministic then there will be a variance in the output which can be measured.
%

\subsection{Definition: Accuracy}

Precision and tolerance are not to be confused with \textit{accuracy} which describes how closely the generated output of a process aligns to a known standard. We therefore define accuracy as the difference between the true value (or reference value) and what has been achieved in the simulation or generation process. 
%
For a driving simulation the reference value, for which accuracy can be measured against, may be the real world that it seeks to emulate, where any divergence from this standard is termed the \textit{reality gap}. In most cases such accuracy will not be possible due to computational demands of such an exact replica and in most cases is unnecessary, where some authors state that `just the right amount of realism' is required to achieve valid simulation results~\cite{Koopman2018}.

\subsection{Definition: Variance Measurement}

If the simulator is non-deterministic then how can this variance be measured?
%
This can be achieved from the actor state, or more specifically from the statistical variance of actor paths within a given scenario with a fixed set of actions. Actor path variance is chosen over other measures as this distance based metric forms the basis for many downstream verification tests such as assertion checking.

The term \textit{simulation variance} refers to a measure of actor state variance in the simulation assuming fixed actions. Results are referred to in terms of deviation, rather than variance, as this is a more intuitive dimension to comprehend when considering verification tests, e.g. vehicle close-passing. Moreover, the maximum observed value of deviation was chosen over a mean value to ensure sufficient sensitivity.

\subsection{Definition: Simulation Trace}

A simulation trace is the output log from the simulator which is a time series of all actor positions ($x,y,z$) at regular intervals based on the time-step $dt$. This definition could be extended to include other variables, but position is sufficient for this analysis.

\subsection{Definition: Scene, Scenario \& Situation}

We will adopt the terminology defined in~\cite{Ulbrich2015}, where \textit{scene} refers to all static objects including the road network, street furniture, environment conditions and a snapshot of any dynamic elements. Dynamic elements are the elements in a scene whose actions or behaviour may change over time; these are considered actors and may include the AV or \textit{ego vehicle}, other road vehicles, cyclists, pedestrians and traffic signals. The \textit{scenario} is then defined as a temporal development between several scenes which may be specified by specific parameters. A \textit{situation} is defined as the subjective conditions and determinants for behaviour at a particular point in time.

@@

% This was calculated by averaging the variance between all of the different runs through out the simulation time for all of the agents, then taking the square root to obtain the total mean devaiton. Note the dotted red line in the plot shows the millimeter (mm) level. Below this level we assume that the simulation variance is low enough that it would not affect our assertion checking for V\&V of AV simulations purposes, that it is sufficient to assume the tests are deterministic and repeatable. This line can be moved up or down and it is entirely up to verification engineers where they want to define the level below which tests' variance is negligibale and it is suffcient to assume that tests are deterministic.
%
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.99\linewidth]{Other/Figures/Methodology_Diagram_v4.pdf}
    \caption{Flow diagram of the methodology proposed to determine the simulation variance of a simulation.}
    \label{method_diagram}
\end{figure*}

@@

\subsection{Future Work}
A potential avenue of exploration is developing a version of the simulator which has been designed specifically for verification. In this version the simulator would be deterministic because the execution is suitably controlled along with sources of randomness and scheduling. A similar task was achieved with the \texttt{RR} program.\footnote{\url{https://github.com/mozilla/rr}}

>>>>

\subsection{Definitions}
\subsubsection{Determinism}
% What is a deterministic system
Schumann describes determinism as the property of causality given a temporal development of events such that any state is completely determined by prior states~\cite{Schumann2010}. However, in the context of simulation this should be expanded to include not just prior states but also the history of actions taken by all actors. Therefore, a deterministic simulation will always produce the same result given the same history of prior states and actions.
% 
% Herein the term \textit{simulation variance} is used to refer to how simulation traces and outputs vary, when the same stimulus is provided repeatedly to the simulation environment under identical initial conditions. Specifically, we focus on the variance of actor paths within a given scenario.

A simulation can be thought of as the generation or production of experimental data. 
%
In the case of a driving simulator, kinematics will describe future states of actors given the current conditions and actions taken, thereby generating new data. 
%
If a simulation is deterministic then there will be no variation in the generated output data, i.e. all future states are perfectly reproducible from prior states and actions. 
%
However, if a simulation is non-deterministic then then there will be a variation in the generated output data. \\

@@

% \subsection{Definitions: Precision, Variance \& Tolerance}
\subsubsection{Precision, Variance \& Tolerance}

@@

If a simulator is deterministic then it will produce results with absolute precision or zero variance and hence will be within acceptable tolerance. If the simulator is non-deterministic then there will be a variance in the output which can be measured.\\
%

% \subsection{Definition: Accuracy}
\subsubsection{Accuracy}
Precision and tolerance are not to be confused with \textit{accuracy} which describes how closely the generated output of a process aligns to a known standard. We therefore define accuracy as the difference between the true value (or reference value) and what has been achieved in the simulation or generation process. 
%
For a driving simulation the reference value, for which accuracy can be measured against, may be the real world that it seeks to emulate, where any divergence from this standard is termed the \textit{reality gap}. In most cases such accuracy will not be possible due to computational demands of such an exact replica and in most cases is unnecessary, where some authors state that `just the right amount of realism' is required to achieve valid simulation results~\cite{Koopman2018}. \\

% \subsection{Definition: Variance Measurement}
\subsubsection{Variance Measurement}
If the simulator is non-deterministic then how can this variance be measured?
%
This can be achieved from the actor state, or more specifically from the statistical variance of actor paths within a given scenario with a fixed set of actions. Actor path variance is chosen over other measures as this distance based metric forms the basis for many downstream verification tests such as assertion checking.

The term \textit{simulation variance} refers to a measure of actor state variance in the simulation assuming fixed actions. Results are referred to in terms of deviation, rather than variance, as this is a more intuitive dimension to comprehend when considering verification tests, e.g. vehicle close-passing. Moreover, the maximum observed value of deviation was chosen over a mean value to ensure sufficient sensitivity.\\

% \subsection{Definition: Simulation Trace}
\subsubsection{Simulation Trace}
A simulation trace is the output log from the simulator which is a time series of all actor positions ($x,y,z$) at regular intervals based on the time-step $dt$. This definition could be extended to include other variables, but position is sufficient for this analysis.\\

% \subsection{Definition: Scene, Scenario \& Situation}
\subsubsection{Scene, Scenario \& Situation}
We will adopt the terminology defined in~\cite{Ulbrich2015}, where \textit{scene} refers to all static objects including the road network, street furniture, environment conditions and a snapshot of any dynamic elements. Dynamic elements are the elements in a scene whose actions or behaviour may change over time; these are considered actors and may include the AV or \textit{ego vehicle}, other road vehicles, cyclists, pedestrians and traffic signals. The \textit{scenario} is then defined as a temporal development between several scenes which may be specified by specific parameters. A \textit{situation} is defined as the subjective conditions and determinants for behaviour at a particular point in time.

@@

% This was calculated by averaging the variance between all of the different runs through out the simulation time for all of the agents, then taking the square root to obtain the total mean devaiton. Note the dotted red line in the plot shows the millimeter (mm) level. Below this level we assume that the simulation variance is low enough that it would not affect our assertion checking for V\&V of AV simulations purposes, that it is sufficient to assume the tests are deterministic and repeatable. This line can be moved up or down and it is entirely up to verification engineers where they want to define the level below which tests' variance is negligibale and it is suffcient to assume that tests are deterministic.
%
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.99\linewidth]{Other/Figures/Methodology_Diagram_v5.pdf}
    \caption{Flow diagram of the methodology proposed to determine the simulation variance of a simulation.}
    \label{method_diagram}
\end{figure*}

@@

% \subsection{Future Work}
A potential avenue of exploration is developing a version of the simulator which has been designed specifically for verification. In this version the simulator would be deterministic because the execution is suitably controlled along with sources of randomness and scheduling. A similar task was achieved with the \texttt{RR} program.\footnote{\url{https://github.com/mozilla/rr}}
